Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-14 docs/ACM/paper.tex

Annotation systems have risen to a place of prominence as a simple and
effective means to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11}, Cilk++~\cite{Leiserson09} and
Tareador~\cite{Ayguade15}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
%\guido{What do you mean by modern accelerators? Not clear.}
Combined with modern hardware accelerators such as GPUs and FPGAs, they have led
to substantial performance gains~\cite{Bertolli14,Martineau16,Mendonca17,Poesia17,Reyes12,Antao16,Wienke12}.
%\guido{Suppport what?}
Nevertheless, although convenient, the use of annotations is not straightforward, 
and still lacks supporting tools that help programmers to check if annotations
are correct and/or effective.
%replaced:
%Yet, even if convenient,

@@

Definition~\ref{def:swe} is too unconstrained: a function $W(G) = 0$ for any $G$
satisfies it.
However, we want $W$ to approximate the dynamic behavior of programs.
The literature contains heuristics to implement $W$.
Perhaps, the best-known among these heuristics is the static profiler of
Wu and Larus~\cite{Wu94}.
In this paper, we adopted a different approach: we reuse the symbolic range
analysis of Section~\ref{sub:sra} to augment static data with dynamic information.
In other words, we use symbolic range analysis to construct expressions that
represent the number of iterations of loops.
A similar approach has already been used to decide when to migrate virtual memory
pages in NUMA architectures~\cite{Piccoli14}.
OpenMP 4.0 contains syntax to enable the conditional creation of tasks.
Our symbolic analysis lets us build predicates for such conditional expressions.

@@

% Explain how we use symbolic information to compute costs.
The declarative rules in Figure~\ref{fig:swe} sketches the heuristics that
we use to compute the SWE for a given hammock region $S$.
Rule \textsc{LoopInf} applies on loops whose number of iterations we cannot bound
via some symbolic expression computed statically.
Rule \textsc{LoopExp} represents loops that we can analyze using our Symbolic
Range Analysis.
The auxiliary function $\mathit{Iter}(S)$ returns a symbolic expression that
represents
the range of values covered by the code $S$ that controls the number of
iterations of the loop.
In Figure~\ref{fig:alg_main}, these rules are implemented by function
\textsf{profitabilityAnalysis}.

\paragraph{The Cost Model.}
A static workload estimate is only meaningful in the context of a {\em cost
model}.
A cost model is a collection of parameters that determine the impact of the
underlying computer architecture onto the creation and management of tasks.
The literature contains examples of techniques used to build cost models,
be it analytically~\cite{Baghsorkhi10}, or empirically~\cite{Poesia17}.
The automatic construction of a cost model is not the goal of this paper.
Instead, the reader shall notice that Figure~\ref{fig:alg_main} receives a
cost model as a parameter.
For the experiments in Section~\ref{sec:eval}, we have determined a small
collection of constants related to the creation of threads in our target
architecture, including the value of \textsf{WORK\_CUTOFF}, seen in
Example~\ref{ex:cond_task}.

@@

The benchmarks used in Section~\ref{sub:performance} have been coded to
demonstrate the power of parallel systems; hence, they have
been written in a way that simplifies the discovery of parallelism in programs.
In this section, we address the following question: ``can \Taskminer" find
parallelism in general programs?
To answer it, we have applied \Taskminer{} onto the 219 C programs available in
the LLVM test suite, and have compared the annotated version
against the sequential version compiled.
\Taskminer{} annotates a benchmark if it can find symbolic bounds to every
memory access (load or store) used within a region marked as a task candidate.
Under this constraint, we have been able to discover tasks in 63 benchmarks.
Figure~\ref{fig:TM_Versatility} relates the number of tasks and the number of
instructions in the 30 largest benchmarks that we analyze.
To avoid counting multiple C files for the same benchmark,
Figure~\ref{fig:TM_Versatility} contains only benchmarks that
consists of a single file (present in LLVM's \textsf{SingleSource} folder).
As this experiment demonstrates, \Taskminer{} {\em can annotate a non-trivial
number of real-world benchmarks}.

>>>>

Annotation systems have risen to a place of prominence as a simple and
effective means to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11}, Cilk++~\cite{Leiserson09} and
OmpSs~\cite{bueno:icpp:2011, duran:ppl:2011}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
%\guido{What do you mean by modern accelerators? Not clear.}
Combined with modern hardware accelerators such as GPUs and FPGAs, they have led
to substantial performance gains~\cite{Bertolli14,Martineau16,Mendonca17,Poesia17,Reyes12,Antao16,Wienke12}.
%\guido{Suppport what?}
Nevertheless, although convenient, the use of annotations is not straightforward, 
and still lacks supporting tools that help programmers to check if annotations
are correct and/or effective.
%replaced:
%Yet, even if convenient,

@@

Definition~\ref{def:swe} is too unconstrained: a function $W(G) = 0$ for any $G$
satisfies it.
However, we want $W$ to approximate the dynamic behavior of programs.
The literature contains heuristics to implement $W$.
Perhaps, the best-known among these heuristics is the static profiler of
Wu and Larus~\cite{Wu94}.
In this paper, we adopted a different approach: we reuse the symbolic range
analysis of Section~\ref{sub:sra} to augment static data with dynamic information.
In other words, we use symbolic range analysis to construct expressions that
represent the number of iterations of loops.
A similar approach has already been used to decide when to migrate virtual memory
pages in NUMA architectures~\cite{Piccoli14}.
OpenMP 4.0 contains syntax to enable the conditional creation of tasks.
Our symbolic analysis lets us build predicates for such conditionals.

@@

% Explain how we use symbolic information to compute costs.
The declarative rules in Figure~\ref{fig:swe} sketch the heuristics that
we use to compute the SWE for a hammock region $S$.
Rule \textsc{LoopInf} applies on loops whose number of iterations we cannot bound
via some symbolic expression computed statically.
We assume that loops execute 10 times, following previous
statistics~\cite{Wu94}.
Rule \textsc{LoopExp} represents loops that we can analyze using our Symbolic
Range Analysis.
The auxiliary function $\mathit{Iter}(S)$ returns a symbolic expression that
represents
the range of values covered by the code $S$ that controls the number of
iterations of the loop.
In Figure~\ref{fig:alg_main}, these rules are implemented by function
\textsf{profitabilityAnalysis}.

\paragraph{The Cost Model.}
A static workload estimate is only meaningful in the context of a {\em cost
model}.
A cost model is a collection of parameters that determine the impact of the
underlying computer architecture onto the creation and management of tasks.
The literature contains examples of techniques used to build cost models,
be it analytically~\cite{Baghsorkhi10}, or empirically~\cite{Poesia17}.
The automatic construction of a cost model is not the goal of this paper.
Instead, the reader shall notice that Figure~\ref{fig:alg_main} receives a
cost model as a parameter.
For the experiments in Section~\ref{sec:eval}, we have determined a small
collection of constants related to the creation of threads in our target
architecture.
As an example, the value of \textsf{WORK\_CUTOFF}, seen in
Example~\ref{ex:cond_task}, in our setting, is roughly 500 cycles.

@@

The benchmarks used in Section~\ref{sub:performance} have been coded to
demonstrate the power of parallel systems; hence, they have
been written in a way that simplifies the discovery of parallelism in programs.
In this section, we address the following question: ``can \Taskminer" find
parallelism in general programs?
To answer it, we have applied \Taskminer{} onto the 219 C programs available in
the LLVM test suite, and have compared the annotated version
against the sequential version compiled.
\Taskminer{} annotates a benchmark if:
(i) it can find symbolic bounds to every
memory access used in a vane; and
(ii) the vane is large enough to pay for the cost of creating threads.
Under these constraints, we have discovered tasks in 63 benchmarks.
Figure~\ref{fig:TM_Versatility} relates the number of tasks and the number of
instructions in the 30 largest benchmarks that we have used.
To avoid counting multiple C files for the same benchmark,
Figure~\ref{fig:TM_Versatility} contains only benchmarks that
consists of a single file (present in LLVM's \textsf{SingleSource} folder).
As this experiment demonstrates, \Taskminer{} {\em can annotate a non-trivial
number of real-world benchmarks}.
